### 2019-08-14 Writeup/Thesis/pages/lit_review.tex
Changes; - More corrections made

The organization of the section is as follows:

@@

The starkest limitations of the literature are the range of participants and encoding entropy. The worst studies tested only 22-bits of entropy. This lack of entropy makes it challenging to compare results directly. One of the papers this most effects is the early work by \textbf{Hsiao \textit{et al.}}\cite{hsiao2009study}  where their highest entropy is 28-bits. This level of entropy was inadequate even at the time of publication. There is an attempt by the authors' to address this issue in the later stages of the paper where they write \textit{``[...] increasing entropy is not a solution because it sacrifices usability and accuracy. With more entropy, representations will contain longer sequences of characters or more minute details which will lead to increased time and errors during comparisons.''}. This statement is backed up with no empirical evidence, and the authors fail to consider how the low entropy would affect the overall security of the schemes.

Attacker strength is also another metric used to compare results concluded by each paper. Some papers failed to address their attacker strength consideration directly, but the overall strength has been inferred from the changes made to their schemes. For example, if they decided to change a single character in a 40-digit (160-bit) SHA-1 hex digest, they are indirectly stating that the attacker can control 39-digits (156-bit). To achieve this, the attacker would have to compute $2^{156}$ SHA-1 compressions to find a key match.
In the literature, this element ranged from $2^{28}$ to around $2^{242}$; however, this has some relation to the size of the encoding schemes used. This is a substantial range that makes it ultimately challenging to compare and confer results confidently. Further work could exclusively look into the affects attacker strength has on the success of attacks. Moreover, all the papers assessed failed to fully consider the feasibility of attacks in terms of computer and storage requirements. This again, shows gaps in the literature.

All of the studies considered demographical data when presenting their results. Their average ages were all around $\sim35$ years old with the majority of participants educated with at least a bachelor degree. This was alongside the equal split between male and female participants.

One consideration of note is that made by Dechand \textit{et al.} \cite{dechand2016empirical} where they briefly consider medical conditions such as ADHD and reading or visual disorders and the way they affect the comparison's effectiveness. They highlight a slight reduction in overall accuracy, although due to their small sample size, they cannot conclusively validate these results. This is an aspect unique to all literature. Further work would be required to produce conclusive results. Therefore, highlighting a gap in the literature.

One glaring issue with the demographical health of \textbf{Ersin Uzun \textit{et al.}}\cite{uzun2007usability} study is the use of two entirely different groups of participants. Not only were their demographics different, but they were from different countries (America and Finland). Different cultures contain inherent biases and assumptions. Moreover, more concerns are raised around the dual study design of this paper with two completely different set of participants. Changes were made pragmatically regarding the results from the first round of 40 participants. These were then re-assessed and the results were directly compared. This puts huge doubts on the validity of the results with no consideration made by the authors to control external factors that may affect the performance of the method of comparison.

Hexadecimal and numerical schemes were not included as encoding schemes in work by \textbf{Hsiao \textit{et al.}}\cite{hsiao2009study} (some of the most widespread encoding schemes). This decision was based on their claims on similarities to Base32 and ``well-known deficiencies" in the excluded schemes. This point was provided with no further justification or quantified in any way. It is also not consistent with available research, for example, in \textit{``Empirical Study of Textual Key-Fingerprint Representations''}\cite{dechand2016empirical} it was shown numerical representations performed significantly better than that of Base32.

@@

Table \ref{tab:attacker} has been provided to visually compare the differing aspects of the papers' parameters. Clearly it can be seen from the table the large ranges in participant size, entropy and attacker strength.

\subsubsection*{Topic conclusion}
Overall, this review has identified several key areas suitable for further work.
The first is the performance assessment of graphical encoding schemes. Recreation of pre-existing results from \cite{hsiao2009study}\cite{tan2017can} is required to corroborate current conclusions and validate results.
Another gap in the research is the consideration into the utilization of encoding schemes in realistic conditions, i.e. ``remove vs. proximity." This topic was initially covered by \cite{shirvanian2017pitfalls} but their scope was limited. Further work, therefore, could increased the scope and touch upon a large number of schemes in these settings.
The final aspect for further work is the limited consideration into the feasibility of attacks on encoding schemes. All of the papers assessed simulated attacks and had minimal consideration for the execution of these attacks. Therefore, further work could delve into the implementation of such attacks and their feasibility in terms of compute and storage complexity.

\subsection{Encoding schemes}
Another area of research is investigations into the actual physical encodings of the hash digest. This section will briefly discuss the current research available on the creation and security of actual encoding schemes. The actual details of the operation of the schemes are outside the scope of this literature review, therefore, minimal attention will be allocated to these details.

Some of the oldest preliminary work into visual encoding schemes was performed by \textbf{Adrian Perrig \textit{et al}}\cite{perrig1999hash}. in the creation of their scheme ``Random Art" in 1999. The motivation for creating such a scheme was the perceived flaws in the ways humans verify and compare written information. As mentioned in previous sections visual encoding schemes have been shown to have mixed success, with low security being one of their most alarming flaws. This research laid the foundation for further work in analysing the security of visual encoding schemes.

Further research into the creation of unique visual hash schemes have been performed by \textbf{C Ellison \textit{et al.}} \cite{ellison2003public} (Flag), \textbf{Yue-Hsun Lin \textit{et al.}}\cite{lin2010spate} (T-Flag) and work by \textbf{M.  Olembo \textit{et al.}}\cite{olembo2013developing}. Each publication has provided a new way to visually represent a key fingerprint. Alongside the academic literature, there are more informally presented methods of visual fingerprints such as Unicorns\footnote{https://unicornify.pictures/} and Robots\footnote{https://github.com/e1ven/Robohash}. This list is by no means exhaustive but is used to depict the amount of research and work invested into graphical hash representations.

One paper of note is the preliminary work performed by \textbf{D Loss \textit{et al.}}\cite{loss2009drunken} in their \textit{``An analysis of the OpenSSH fingerprint visualization algorithm''} where their aim was to spur on further research with their initial findings into the security of the OpenSSH scheme. The authors claim that the use of the algorithm in OpenSSH is only heuristically defined and there is a need for a formal proof of its security. \\
The paper proposed a number of ways to generate similar fingerprints. The methods proposed were: Naive brute force, Graph Theory, and brute force of a full visual set. They were only able to produce only very basic results and have proposed a large amount of potential further work. Since the paper's publication in 2009, there seems to have been no research building on the work of the authors. This highlights a current gap in the available literature.

Minimal research has also focused on basic textual fingerprint representations and their respective security. Work by \textbf{A. Karole} and \textbf{N. Saxena}\cite{karole2009improving} looked into ways to improve the security of a textual representation. This research aim was to improve the secure device pairing process of comparing two numerical values. The devices used (Nokia 6030b; Mid-range devices at the time of publication) and the SAS compared results in findings that are not directly applicable in a fingerprint comparison context.

A more specific subsection of textual fingerprints is the use of words and sentences to encode hash digests. Some of the first work in this area was produced by \textbf{Juola} and \textbf{Zimmermann}
\cite{juola1996whole} and their work in generating a word list where phonetic distinctiveness was prioritised. Each word is mapped to a single byte. The unique aspect of the word list is the separation of ``even'' and ``odd'' words where ``even'' byte positions are sample from the even-list and ``odd'' from the odd-list. This effectively creates two sub-word lists. The maximisation of linguistic distinctiveness of these word lists were maximised through the use of a Genetic algorithm. The paper also includes a study on effective measures of ``linguistic distances'' of words and provided an in-depth discussion into these areas.

@@

Other research of note is work by \textbf{M. Goodrich \textit{et al.}}\cite{goodrich2006loud} called \textit{Loud and Clear: Human-Verifiable Authentication Based on Audio}. As the name suggests the authors were researching ways to improve current methods of secure device pairing. The unique aspect of this work is the use of a Text-to-Speech system reading out syntactically correct English sentences. The sentences are based on a MadLibs\footnote{https://en.wikipedia.org/wiki/Mad\_Libs} where static placeholders were replaced with potential words.\\
The work into a potential wordlist can be seen as an extension to the work performed by Juola and Zimmermann\cite{juola1996whole} as they aimed to emulate the techniques used in PGPfone. The paper's finding are limited by the lack of empirical data backing up claims made by the author as the systems performance and security are only theoretically assessed.

Aside from this research, there have been further informal implementations of fingerprint encodings. The first being by \textbf{Michael Rogers}\footnote{https://github.com/akwizgran/basic-english}. Rogers' implementation is a program designed to map fingerprints to pseudo-random poems. This implementation was again, empirically evaluated by Dechand \textit{et al.}\cite{dechand2016empirical}. Older work by \textbf{N. Haller} with the S/KEY\cite{haller1995s} shows the implementation of a system designed to represent a hash as a series of six short words. However, this system is designed for a one-time-password purpose and only provides word mappings for basic human usability of the password and not within a fingerprint verification context. Therefore, the wordlist has not been designed with pronounceability in mind.

\subsubsection*{Pretty Easy Privacy}
\label{sec:pep}
A very recent implementation of a word list can be found in Pretty Easy Privacy (\pep) implementation of TrustWords\footnote{https://tools.ietf.org/html/draft-birk-pep-trustwords-03}. \pep is a data encryption system that utilises PGP encryption to provide E2EE on all common channels of communication such as email or SMS. The imbedded design principles state that above all the systems should be easy to install, use and understand.

@@

\pep deals with the threat of MiTM attacks by having users compare the respective key fingerprints encoded as a set of words. Figure \ref{fig:trustwords} shows the \pep Android implementation of Trustwords. The users then authenticate the words on an OOB (Out of Band) channel such as a phone call or in-person communication. If both users decide the words match they will accept or decline respectively.

The unique aspect of TrustWords is its mapping of a single word to 16-bits. In comparison to other literature, this is the highest number of bits-per-word seen. Full mappings (no duplication of words) would, therefore, require $2^{16}$ words in the dictionary where arguably the dictionary size is higher than most users' vocabulary. This deviation from the norm has not been currently backed up by research.

@@

Alongside the abnormally high number of words the design choice to exclude slang and profanities from the English Trustword dictionary requires dual-mapping of a section of words. Approximately 13633/65536 (20.8\%) of words are re-mapped in the dictionary leaving 51903 unique words. The re-mapping is also done on a loop with it remaining alphabetical. Figure \ref{fig:remap} shows the position in the dictionary where this occurs. This predictability within the dictionary will be explored later in the paper.

Moreover the main RFC documentation still remains in a draft stage and states \textit{``It is for further study, what minimal number of words (or entropy) should be required.''}. These aspects clearly highlight on a gap in the current literature.

@@

In conclusion to this topic, the current research has primarily focused on the research and creation of visual representations. Research for textual fingerprints is fragmented and incomplete with work Juola and Zimmermann 
\cite{juola1996whole} and M. Goodrich \textit{et al.}\cite{goodrich2006loud} providing meaningful research to build upon in terms of word a sentence based encodings. The fragmentation of this research leaves room for further work into this topic area. Alongside this, findings from the previous sections research shows that human language based encodings provided the best usability and, therefore, should be a target for further research looking to improve upon their security and usability.

\subsection{Attacks on encoding schemes}
This area of research studies ways to physically execute attacks on fingerprint encoding schemes. This differs from previously examined work due to papers discussing the performance and fallibility of encoding schemes simulated the attack without consideration for how the attack would be performed. Research in this area is scant, with lots of research attention being directed towards the security of Man-in-the-Middle (MITM) attacks and not the encoding schemes themselves.

Research in 2002 by \textbf{Konrad Rieck}\cite{rieck2002fuzzy} is the first formalisation of attacks on fingerprint representations. The paper titled \textit{``Fuzzy Fingerprints Attacking Vulnerabilities in the Human Brain''}
aimed to look into ways users check hexadecimal encoded OpenSSH fingerprint representations. The author created an elegant way to `weight' more important chunks of the digest. The bytes furthest to the right and left of the digests provided the highest weight. The weight was the smallest in the centre of the digest. This provides a way to score digests and determine the best partial collisions found. For example with the target fingerprint: \verb|9F23| a partial match \verb|9313| is given a score of 45\% even though only two characters were matching. This is due to the weightings.

The paper contains an implementation with a ``1.2GHz CPU'' being able to obtain 130,000 H/s (With MD5). In comparison to this, a mid-range Intel i5-3320M CPU can today obtain 111,700,000 MD5 H/s. This shows that the results obtained from the paper are significantly outdated. However, even with the low hash rate, the author was able to obtain some promising results. Figure~\ref{ref:fuzz} contains the best example used.

@@

Overall the paper shows an interesting way to create partial fingerprint matches but is not quantified by any empirical evidence gathered on real world users. This, therefore, highlights on gaps in the coverage of this literature.

The only other relevant research on this topic is the work by \textbf{M Shirvanian \textit{et al.}}\cite{shirvanian2014wiretapping} 
and their paper \textit{``Wiretapping via Mimicry: Short 
Voice Imitation Man-in-the-Middle Attacks on Crypto 
Phones''}. Further research in the area of ``human voice impersonation'' has received lots of attention \cite{mukhopadhyay2015all}\cite{chen2017you}\cite{wu2015spoofing}. This paper was chosen over other alternatives due to is specific use of encoding schemes in its evaluation.

In this paper, the authors develop a way to 
impersonate users when authenticating 
Short-authentication-Strings (SAS) in pairing of 
Crypto-phones. To achieve this impersonating they propose 
two methods: ``Short voice reordering attack'' where an 
arbitrary SAS string is recreated by re-ordering snipped 
obtained from eavesdropping a previous connection
and ``Short voice morphing attacks'' whereby the use of 
previously eavesdropped audio snippets the attacker can
morph their own voice to match that of the victim. With 
these methods, they aimed to attack encodings of Numbers, 
PGP word list (previously discussed work by Juola and 
Zimmermann \cite{juola1996whole}) and MadLib (M. Goodrich 
\textit{et al.}\cite{goodrich2006loud} work also 
previously discussed). The effectiveness of these attacks 
were evaluated with a study involving 30 participants.

Results from the paper show the effectiveness of these 
methods. Compared to the baseline of the attacker's voice 
replacing the victim where this performed with a 
$\sim$18\% success rate. Morphing gained an 
overall success rate of 50.58\% and Reordering a very 
impressive 78.23\% success rate. Showing that these 
attacks provide an improvement on top of the naive implementation.

One of the biggest limitations addressed by the authors 
was the reduction in success rates as the size of the 
authentication string grew. The morphing and reordering 
attacks become increasingly ineffective as the user has 
more time to detect imperfections. This is not quantified by 
the author and the extent of this degradation is never 
empirically discussed. Therefore, the results from this 
study are only effective and applicable in a SAS context.

@@

\section{Overall Summary}
In summary of the literature gaps areas discovered; encoding scheme performance requires further work in the performance of all graphical schemes to back up the results made by other work. Furthermore, there is a gap with the assessment of encoding schemes in the context of "remote-vs-proximity" first proposed by M. Shirvanian et al. as it would arguably provide a better simulation of real world scenarios.

In the context of human attributes and the way they alter the performance of the schemes; further research is required into how the fluency of languages affects authentication ceremony. This is a continuation of the initial work by Hsiao et al and could allow the creation of schemes that are better suited to certain languages.
Alongside this, there is a research gap in the investigation of how mental impairments affect the performance. This is useful due to the number of potential users being affected due to systems not being designed with them in mind. For example, 7\% of the population has been identified as having dyslexic tendencies\cite{peterson2012developmental}. This means with a UK population of around 63,000,000\footnote{From the 2011 Census collated by the Office for National Statistics}  4,410,000 people could benefit from encoding schemes including dyslexic oriented benefits. This is also only one of many visual impairments that could affect a scheme's performance, thus further contributing to the substantial affect further work could have on this area.

Other possible large areas for consideration is the lacks of direct consideration for attack strength when assessing the vulnerability of encoding schemes. Issues included the wide range of attack strength ($2^{28}$ - $2^{242}$) and the lack of consideration at all from certain papers. This alongside the complete lack of computation and storage complexity considerations means this is a prime area for further research as it would improve the applicability of results.

The final major research gap identified is the lack of justification into the newly created \pep Trustwords. Abnormal design choices (16-bit per word) and its recent creation makes it another area for a wide variety of further research. This is also combined with the conclusive results from the literature on the effectiveness of language based encodings such as words. This would suggest that research into the improvement of Trustword encoding would be highly beneficial.

>>>>

The organisation of the section is as follows:

@@

The starkest limitations of the literature are the range of participants and encoding entropy. The worst studies tested only 22-bits of entropy. This lack of entropy makes it challenging to compare results directly. One of the papers this most effects is the early work by \textbf{Hsiao \textit{et al.}}\cite{hsiao2009study}  where their highest entropy is 28-bits. This level of entropy was inadequate even at the time of publication. There is an attempt by the authors' to address this issue in the later stages of the paper where they write: \textit{``[...] increasing entropy is not a solution because it sacrifices usability and accuracy. With more entropy, representations will contain longer sequences of characters or more minute details which will lead to increased time and errors during comparisons.''}. This statement is backed up with no empirical evidence, and the authors fail to consider how the low entropy would affect the overall security of the schemes.

Attacker strength is also another metric used to compare results concluded by each paper. Some papers failed to address their attacker strength consideration directly, but the overall strength has been inferred from the changes made to their schemes. For example, if they decided to change a single character in a 40-digit (160-bit) SHA-1 hex digest, they are indirectly stating that the attacker can control 39-digits (156-bit). To achieve this, the attacker would have to compute $2^{156}$ SHA-1 compressions to find a key match.
In the literature, this element ranged from $2^{28}$ to around $2^{242}$. However, this has some relation to the size of the encoding schemes used. This substantial range makes it challenging to compare and confer results confidently. 
Further work could exclusively look into the effects that attacker strength has on the success of attacks. Moreover, all the papers assessed failed to fully consider the feasibility of attacks in terms of computer and storage requirements.

All of the studies considered demographical data when presenting their results. Their average ages were all around $\sim35$ years old with the majority of participants educated with at least a bachelor degree. Alongside, the equal split between male and female participants.

One consideration of note is that made by Dechand \textit{et al.} \cite{dechand2016empirical} where they briefly consider medical conditions such as ADHD and reading or visual disorders and the way they affect the comparison's effectiveness. The results highlight a slight reduction in overall accuracy, however, due to their small sample size, the results cannot be considered conclusive. This aspect is unique to all literature assessed. Further work would be required to produce conclusive results. Therefore, highlighting a gap in the literature.

One glaring issue with the demographical health of \textbf{Ersin Uzun \textit{et al.}}\cite{uzun2007usability} study is the use of two entirely different groups of participants. Not only were their demographics different, but they were from different countries (America and Finland). Different cultures contain inherent biases and assumptions. Changes were made pragmatically regarding the results from the first round of 40 participants. These were then re-assessed, alongside the direct comparison of results. This aspect casts vast doubts on the validity of the results with no consideration made by the authors to control external factors that may affect the performance of the method of comparison.

Hexadecimal and numerical schemes were not included as encoding schemes in work by \textbf{Hsiao \textit{et al.}}\cite{hsiao2009study} (some of the most widespread encoding schemes). The justification was the schemes similarities to Base32  alongside "well-known deficiencies". This point was provided with no further justification or quantified. Moreover, it is also inconsistent with available research, for example, in \textit{``Empirical Study of Textual Key-Fingerprint Representations''}\cite{dechand2016empirical} it was shown numerical representations performed significantly better than that of Base32.

@@

Table \ref{tab:attacker} has been provided to compare the different aspects of the papers' parameters visually. It can be seen from the table the substantial ranges in participant size, entropy and attacker strength.

\subsubsection*{Topic conclusion}
Overall, this review has identified several key areas suitable for further work.
The first is the performance assessment of graphical encoding schemes. Recreation of pre-existing results from \cite{hsiao2009study}\cite{tan2017can} is required to corroborate current conclusions and validate results.
Another gap in the research is the consideration into the utilisation of encoding schemes in realistic conditions, i.e. ``remove vs proximity." This topic was initially covered by \cite{shirvanian2017pitfalls}, but their scope was limited. Further work, therefore, could increase the scope and touch upon a large number of schemes in these settings. The final aspect for further work is the limited consideration into the feasibility of attacks on encoding schemes. All of the papers assessed simulated attacks and had minimal consideration for the execution of these attacks. Therefore, further work could delve
into the implementation and feasibility of such attacks.

\subsection{Encoding schemes}
Another area of research is investigations into the actual physical encodings of the hash digest. This section discusses the current research on the creation and security of actual encoding schemes. The technical details of the operation of the schemes are outside the scope of this literature review. Therefore, minimal attention is allocated to these details.

Some of the oldest preliminary work into visual encoding schemes is by \textbf{Adrian Perrig \textit{et al.}}\cite{perrig1999hash} in the creation of their scheme ``Random Art" in 1999. The motivation for creating such a scheme was the perceived flaws in the ways humans verify and compare written information. As mentioned in previous sections, visual encoding schemes have been shown to have mixed success, with low security being one of their most alarming flaws. This research laid the foundation for further work in analysing the security of visual encoding schemes.

Further research into the creation of unique visual hash schemes has been performed by \textbf{C Ellison \textit{et al.}} \cite{ellison2003public} (Flag), \textbf{Yue-Hsun Lin \textit{et al.}}\cite{lin2010spate} (T-Flag) and work by \textbf{M.  Olembo \textit{et al.}}\cite{olembo2013developing}. Each publication has provided a new way  to represent a key fingerprint visually. Alongside the academic literature, there are informally presented methods of visual fingerprints such as Unicorns\footnote{https://unicornify.pictures/} and Robots\footnote{https://github.com/e1ven/Robohash}. This list is by no means exhaustive but is used to depict the amount of research and work invested into graphical hash representations.

One paper of note is the preliminary work performed by \textbf{D Loss \textit{et al.}}\cite{loss2009drunken} in their \textit{``An analysis of the OpenSSH fingerprint visualisation algorithm''} where  they aimed to spur on further research with their initial findings into the security of the OpenSSH scheme. The authors claim that the use of the algorithm in OpenSSH is only heuristically justified, and a more formal proof is required.\\
The paper proposed several ways to generate similar fingerprints. The methods proposed were: Naive brute force, Graph Theory, and brute force of a full visual set. They were only able to produce only initial results and have proposed a large amount of potential further work. Since the paper's publication in 2009, there seems to have been no research building on the work of the authors.

Minimal research has also focused on elemental textual fingerprint representations and their security. Work by \textbf{A. Karole} and \textbf{N. Saxena}\cite{karole2009improving} looked into ways to improve the security of a textual representation. This research aim was to improve the secure device pairing process of comparing two numerical values. The devices used (Nokia 6030b; Mid-range devices at the time of publication) and the SAS (Short Authentication String) compared results in findings that are not directly applicable in a fingerprint comparison context.

A more specific subsection of textual fingerprints is the use of words and sentences to encode hash digests. Some of the first work in this area was by \textbf{Juola} and \textbf{Zimmermann} \cite{juola1996whole}. Their work aimed to produce a word list where phonetic distinctiveness was prioritised. Each word is mapped to a single byte. The unique aspect of the word list is the separation of ``even'' and ``odd'' words where ``even'' byte indices are sample from the even-list and ``odd'' from the odd-list. This technique effectively creates two sub-word lists.  A genetic algorithm was used to maximise linguistic distance. The paper also includes a study on effective measures of ``linguistic distances'' of words and provided an in-depth discussion into these areas.

@@

Other research of note is work by \textbf{M. Goodrich \textit{et al.}}\cite{goodrich2006loud} called \textit{Loud and Clear: Human-Verifiable Authentication Based on Audio}. As the name suggests, the authors were researching ways to improve current methods of secure device pairing. The unique aspect of this work is the use of a Text-to-Speech system reading out syntactically correct English sentences. MadLibs\footnote{https://en.wikipedia.org/wiki/Mad\_Libs} was the base for these sentences. MadLibs is where static placeholders are replaced with a list potential words.\\
This work is an extension to work performed by Juola and Zimmermann\cite{juola1996whole} as they aimed to emulate the techniques used in PGPfone. The paper's findings are limited by the lack of empirical data backing up claims made by the author the performance of the system, and security is only theoretically assessed.

Aside from this research, there have been further informal implementations of fingerprint encodings. The first being by \textbf{Michael Rogers}\footnote{https://github.com/akwizgran/basic-english}. Rogers' implementation is a program designed to map fingerprints to pseudo-random poems. This implementation was again, empirically evaluated by Dechand \textit{et al.}\cite{dechand2016empirical}. Ealier work by \textbf{N. Haller} with the S/KEY\cite{haller1995s} shows the implementation of a system designed to represent a hash as a series of six short words. However, this system is designed for a one-time-password purpose and only provides word mappings for basic human usability of the password and not within a fingerprint verification context. Therefore, the wordlist has not been designed with pronounceability in mind.

\subsubsection*{Pretty Easy Privacy}
\label{sec:pep}
A very recent implementation of a word list can be found in Pretty Easy Privacy (\pep) implementation of TrustWords\footnote{https://tools.ietf.org/html/draft-birk-pep-trustwords-03}. \pep is a data encryption system that utilises PGP encryption to provide E2EE on all common channels of communication such as email or SMS. The embedded design principles state that above all the systems should be easy to install, use and understand.

@@

\pep deals with the threat of MiTM attacks by having users compare the respective key fingerprints encoded as a set of words. Figure \ref{fig:trustwords} shows the \pep Android implementation of Trustwords. The users then authenticate the words on an OOB (Out of Band) channel such as a phone call or in-person communication. If both users decide the words match, they accept or decline respectively.

The unique aspect of TrustWords is its mapping of a single word to 16-bits. If compared to alternative literature, this is the highest number of bits-per-word seen. Full mappings (no duplication of words) would, therefore, require $2^{16}$ words in the dictionary. This number of words is arguably higher than most users' vocabulary. This deviation from the norm has not been currently backed up by research.

@@

Alongside the abnormally high number of words, the design choice to exclude slang and profanities from the English Trustword dictionary requires dual-mapping of a section of words. Approximately 13633/65536 (20.8\%) of words are re-mapped in the dictionary, leaving 51903 unique words. The re-mapping looped with the dictionary remaining alphabetical. Figure \ref{fig:remap} shows the position in the dictionary where this occurs.

Moreover, the primary RFC documentation remains in a draft stage and states: \textit{``It is for further study, what minimal number of words (or entropy) should be required.''}. These aspects highlight a substantial gap in the current literature.

@@

In conclusion to this topic, the current research has primarily focused on the research and creation of visual representations. Research for textual fingerprints is fragmented and incomplete with work Juola and Zimmermann 
\cite{juola1996whole} and M. Goodrich \textit{et al.}\cite{goodrich2006loud} providing meaningful research to build upon in terms of word and sentence based encodings. The fragmentation of this research leaves room for further work into this topic area. Alongside this, findings from the previous section's research show that human language based encodings provided the best usability and, therefore, should be a target for further research looking to improve upon their security and usability.

\subsection{Attacks on encoding schemes}
This area of research studies ways to physically execute attacks on fingerprint encoding schemes. This topic differs from previously examined work due to papers discussing the performance and fallibility of encoding schemes simulated the attack without consideration for how the attack execution. Research in this area is scant, with lots of research attention on security of Man-in-the-Middle (MITM) attacks and not the encoding schemes themselves.

Research in 2002 by \textbf{Konrad Rieck}\cite{rieck2002fuzzy} is one the first formalisation of attacks on fingerprint representations. The paper titled \textit{``Fuzzy Fingerprints Attacking Vulnerabilities in the Human Brain''} aimed to look into ways users check hexadecimal encoded OpenSSH fingerprint representations. The author created an elegant way to `weight' essential chunks of the digest. The bytes furthest to the right and left of the digests provided the highest weight. This technique provides a way to score digests and determine the best partial collisions found. For example, with the target fingerprint: \verb|9F23| and partial match \verb|9313| is given a score of 45\% even though only two characters are matching.

The paper contains an implementation with a ``1.2GHz CPU'' being able to obtain 130,000 H/s (With MD5). In comparison to this, a mid-range Intel i5-3320M CPU can today obtain 111,700,000 MD5 H/s. This imbalance shows that the results obtained from the paper are significantly outdated. However, even with the low hash rate, the author was able to obtain some promising results. Figure~\ref{ref:fuzz} contains the best example used.

@@

Overall the paper shows an interesting way to create partial fingerprint matches but is not quantified by any empirical evidence gathered on real-world users.

The only other relevant research on this topic is the work by \textbf{M Shirvanian \textit{et al.}}\cite{shirvanian2014wiretapping} 
and their paper: \textit{``Wiretapping via Mimicry: Short 
Voice Imitation Man-in-the-Middle Attacks on Crypto 
Phones''}. Further research in the area of ``human voice impersonation'' has received lots of attention \cite{mukhopadhyay2015all}\cite{chen2017you}\cite{wu2015spoofing}. This paper was chosen over other alternatives due to its specific use of encoding schemes in its evaluation.

In this paper, the authors develop a way to 
impersonate users when authenticating 
Short-authentication-Strings (SAS) in the pairing of 
Crypto-phones. To achieve this impersonating they propose 
two methods: ``Short voice re-ordering attack'' where an 
arbitrary SAS string is recreated by re-ordering snipped 
obtained from eavesdropping a previous connection
and ``Short voice morphing attacks'' whereby the use of 
previously eavesdropped audio snippets the attacker can
morph their voice to match that of the victim. With 
these methods, they aimed to attack encodings of Numbers, 
PGP word list (previously discussed work by Juola and 
Zimmermann \cite{juola1996whole}) and MadLib (M. Goodrich 
\textit{et al.}\cite{goodrich2006loud} work also 
previously discussed). The effectiveness of these attacks 
was evaluated with a study involving 30 participants.

Results from the paper show the effectiveness of these 
methods. Compared to the baseline of the attacker's voice 
replacing the victim, performed with an $\sim$18\% success rate. Morphing gained an overall success rate of 50.58\% and re-ordering a very impressive 78.23\% success rate and showing that these attacks provide an improvement on top of the naive implementation.

One of the most significant limitations addressed by the authors  was the reduction in success rates as the size of the authentication string grew. The morphing and re-ordering  attacks become increasingly ineffective as the user had more time to detect imperfections. This aspect is not quantified by the author and the extent of this degradation is never empirically discussed. Therefore, the results from this study are only effective and applicable in a SAS context.

@@

\section{Overall Summary}
In summary of the literature gaps areas discovered; encoding scheme performance requires further work in the performance of all graphical schemes to back up the results made by other work. Furthermore, there is a gap with the assessment of encoding schemes in the context of "remote-vs-proximity" first proposed by M. Shirvanian et al. as it would arguably provide a better simulation of real-world scenarios.

In the context of human demographics and the way they alter the performance of the schemes; further research is required into how the fluency of languages affects authentication ceremony. This further work would be a continuation of the initial work by Hsiao et al. and could allow the creation of schemes that are better suited to specific languages.
Alongside this, there is a research gap in the investigation of how mental impairments affect the performance. This aspect is useful due to the number of potential users being affected due to systems not being designed with them in mind. For example, 7\% of the population have been identified as having dyslexic tendencies\cite{peterson2012developmental}. This fact means that with a UK population of around 63,000,000\footnote{From the 2011 Census collated by the Office for National Statistics}  4,410,000 people could benefit from encoding schemes designed to work well with dyslexia. Dyslexia, however, is only one of many impairments that could affect a scheme's performance, thus further contributing to the substantial effect further work could have on this area.

Other possible large areas for consideration is the lacks of direct consideration for attack strength when assessing the vulnerability of encoding schemes. Issues included the previously discussed wide range of attack strength ($2^{28}$ - $2^{242}$) alongside the lack of consideration at all from the majority of papers. Moreover, the complete lack of computation and storage complexity considerations means this is a prime area for further research as it would improve the applicability of results.

The final significant research gap identified is the lack of justification for the newly created Trustwords. Abnormal design choices (16-bit per word) and its recent creation alongside the shown effectiveness of language-based encodings, such as words makes this a promising area for further research.

### 2019-08-14 Writeup/Thesis/pages/research_question.tex
Changes; - More corrections made

Following on from the previous chapter, this chapter will define the selected gaps and the subsequent research questions and aims.

The chosen project aims to asses the security of \pep's minimum recommend number of four Trustwords (As stated in Figure \ref{fig:trustwordsNum}). As discussed in the previous chapter, Trustwords aims to sacrifice security for increased usability. The encoding scheme has been designed to assist this by having  the user compare a reduced set of words. Moreover, issues with the dictionary's design such as the presence of homophones and dual-mapping of words shows the potential for possible vulnerabilities.

As discussed in the previous chapter, the identified research gap regarding the minimal consideration of attack complexity would be a suitable addition for this project. This is due to it providing the ability to asses actual performance and, thus, the actual real-world strength of Trustwords while providing research to a scant area. Therefore, the security assessment will be supplemented with complexity considerations and an actual implementation of the attack proposed.

@@

By sampling the \pep documentation they have defined the use case of the Trustword handshake: \textit{A handshake is done by comparing the Trustwords between two users through a separate communication channel (e.g. in person or by phone).}\footnote{https://www.pep.security/docs/general\_information.html\#handshake}. It can be assumed due the increased globalization of the planet that the most common handshake occurrence will be over the phone, and, therefore, will not be in person. This is, therefore, the assumed context the handshake will be occurring in. This means the similarity of words will need to be quantified phonetically instead of visually.

@@

With the previously discussed points in mind the following research questions have been proposed.

@@

    \item What is the time and computation complexity required to generate a 'similar' keys for a targeted key pair? \label{goal:complexity}

@@

With these research questions defined, the project will now discuss the design of the proposed attack.

>>>>

This chapter defines the selected gaps and the subsequent research questions extracted.

The chosen project aims to asses the security of \pep's minimum recommendation of four Trustwords (As stated in Figure \ref{fig:trustwordsNum}). As discussed in the previous chapter, Trustwords aims to sacrifice security for increased usability. The encoding scheme has been designed to assist this by having  the user compare a reduced set of words. Moreover, issues with the dictionary's design, such as the presence of homophones and dual-mapping of words show the potential for possible vulnerabilities.

As discussed in the previous chapter, the identified research gap regarding the minimal consideration of attack complexity would be a suitable addition for this project. This choice is due to it providing the ability to asses actual performance and, thus, the actual real-world strength of Trustwords while providing research to a scant area. Therefore, the security assessment is supplemented with complexity considerations alongside an actual implementation of the attack proposed.

@@

By sampling the \pep documentation they have defined the use case of the Trustword handshake: \textit{``A handshake is done by comparing the Trustwords between two users through a separate communication channel (e.g. in person or by phone)''.}\footnote{https://www.pep.security/docs/general\_information.html\#handshake}. It can be assumed due to the increased globalizatiohat is the time andn of the planet that the most common handshake occurrence is over the phone, and, therefore, not in person. This is the assumed context of the handshake. This means, to create near-matches, the similarity of words is quantified phonetically instead of visually.

@@

With the previously discussed points in mind, the following research questions have been proposed.

@@

    \item What are the time and computation complexity required to generate a 'similar' keys for a targeted key pair? \label{goal:complexity}

@@
